---
title: "Telco Customer Churn Prediction"
author: "Seyoung Jung"
date: "12/11/2018"
output: rmarkdown::github_document
---


```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "README_figs/README-"
)
```




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***



# 1 Objective

Predict behavior to retain customers. You can analyze all relevant customer 
data and develop focused customer retention programs.

***


#2 Data

Each row represents a customer, each column contains customer's attributes described on the column Metadata. 
The raw data contains 7043 rows (customers) and 21 columns (features). The "Churn" column is our target.
Description
- **customerID:** Customer ID
- **gender:** Customer gender (female, male)
- **SeniorCitizen:** Whether the customer is a senior citizen or not (1,0)
- **Partner:** Whether the customer has a partner or not (Yes, No)
- **Dependents:** Whether the customer has dependents or not (Yes, No)
- **tenure:** Number of months the customer has stayed with the company
- **PhoneService:** Whether the customer has a phone service or not (Yes, No)
- **MultipleLines:** Whether the customer has multiple lines or not (Yes, No, No phone service)
- **InternetService:** Customer's internet service provider (DSL, Fiber optic, No)
- **OnlineSecurity:** Whether the customer has online security or not (Yes, No, No internet service)
- **OnlineBackup:** Whether the customer has online backup or not (Yes, No, No internet service)
- **DeviceProtection:** Whether the customer has device protection or not (Yes, No, No internet service)
- **TechSupport:** Whether the customer has tech support or not (Yes, No, No internet service)
- **StreamingTV:** Whether the customer has streaming TV or not (Yes, No, No internet service)
- **StreamingMovies:** Whether the customer has streaming movies or not (Yes, No, No internet service)
- **Contract:** The contract term of the customer (Month-to-month, One year, Two year)
- **PaperlessBilling:** Whether the customer has paperless billing or not (Yes, No)
- **PaymentMethod:** The customer's payment method (Electronic check, Mailed check, Bank transfer (automatic), Credit card (automatic))
- **MonthlyCharges:** The amount charged to the customer monthly
- **TotalCharges:** The total amount charged to the customer
- **Churn:** Whether the customer churned or not (Yes or No)


***

# 3 Packages

```{r message=FALSE, results="hide"}
library(readr)      # read_csv
library(dplyr)      # glimpse, 
library(tidyr)
library(caTools)    # sample.split 
library(ggplot2)    # ggplot
library(reshape2)   # melt
library(gridExtra)  # grid.arrange
library(corrplot)   # corrplot
library(MASS)       # stepAIC
```


***


# 4 Data Cleaning

Load the dataset into R and get a glimpse of our data. 
```{r message=FALSE}
telcodata <- read_csv("telcodata.csv")
glimpse(telcodata)
```

Check if there are missing data. We will delete the corresponding rows.
```{r}
MVinfo <- apply(is.na(telcodata), 2, which)     # 11 missing data
withoutMV <- telcodata[-MVinfo$TotalCharges, ]
```

Coerce all the variables to factor variables except for customerID, SeniorCitizen, tenure, MonthlyCharges, and TotalCharges. 
```{r}
cols <- c(1, 3, 6, 19, 20)
withoutMV[,-cols] <- data.frame(apply(withoutMV[-cols], 2, as.factor))
```

Coerce the SeniorCitizen variable to a factor. 1: Yes, 0: No. 
```{r}
withoutMV <- withoutMV %>% mutate(SeniorCitizen = ifelse(SeniorCitizen == 0, "No", "Yes"))
withoutMV$SeniorCitizen <- as.factor(withoutMV$SeniorCitizen)
```

Get rid of the customerID variable
```{r}
cleandata <- withoutMV[, 2:21]
```

Create a new column that classifies the elements by years, and get rid of tenure and TotalCharges.
```{r}
cleandata <- mutate(cleandata, Year=ifelse(tenure %in% 1:12, "0-1", 
                                           ifelse(tenure %in% 13:24, "1-2", 
                                                  ifelse(tenure %in% 25:36, "2-3", 
                                                         ifelse(tenure %in% 37:48, "3-4", 
                                                                ifelse(tenure %in% 49:60, "4-5", 
                                                                       ifelse(tenure %in% 61:72, "5-6", "6-7")))))))
cleandata$Year <- as.factor(cleandata$Year)
cleandata$tenure <- NULL
cleandata$TotalCharges <- NULL
```


***

# 5 Exploratory Data Analysis

EDA Barplot of Gender vs. Churn
```{r}
subDF1 <- data.frame(cleandata$gender, cleandata$Churn)
Category_Gender <- ggplot(subDF1, aes(cleandata$gender, (..count..)/sum(..count..))) + 
     geom_bar(aes(fill=cleandata$Churn), position="dodge") + 
     theme(plot.title = element_text(hjust = 0.5))
print(Category_Gender + labs(y="Proportion", x="Gender", title="Barplot of Gender vs. Churn") +
           scale_fill_discrete(name="Churn"))
```           
EDA Barplot of SeniorCitizen vs. Churn
```{r}
subDF2 <- data.frame(cleandata$SeniorCitizen, cleandata$Churn)
Category_SeniorCitizen <- ggplot(subDF2, aes(cleandata$SeniorCitizen, (..count..)/sum(..count..))) + 
     geom_bar(aes(fill=cleandata$Churn), position="dodge") + 
     theme(plot.title = element_text(hjust = 0.5))
print(Category_SeniorCitizen + labs(y="Proportion", x="Senior Citizen", title="Barplot of Senior Citizen vs. Churn") +
           scale_fill_discrete(name="Churn"))
```
EDA Barplot of Partner vs. Churn
```{r}
subDF3 <- data.frame(cleandata$Partner, cleandata$Churn)
Category_Partner <- ggplot(subDF3, aes(cleandata$Partner, (..count..)/sum(..count..))) + 
     geom_bar(aes(fill=cleandata$Churn), position="dodge") + 
     theme(plot.title = element_text(hjust = 0.5))
print(Category_Partner + labs(y="Proportion", x="Partner", title="Barplot of Partner vs. Churn") +
           scale_fill_discrete(name="Churn"))
```
EDA Barplot of Dependents vs. Churn
```{r}
subDF4 <- data.frame(cleandata$Dependents, cleandata$Churn)
Category_Dependents <- ggplot(subDF4, aes(cleandata$Dependents, (..count..)/sum(..count..))) + 
     geom_bar(aes(fill=cleandata$Churn), position="dodge") + 
     theme(plot.title = element_text(hjust = 0.5))
print(Category_Dependents + labs(y="Proportion", x="Dependents", title="Barplot of Dependents vs. Churn") +
           scale_fill_discrete(name="Churn"))
```
EDA Barplot of PhoneService vs. Churn
```{r}
subDF5 <- data.frame(cleandata$PhoneService, cleandata$Churn)
Category_PhoneService <- ggplot(subDF5, aes(cleandata$PhoneService, (..count..)/sum(..count..))) + 
     geom_bar(aes(fill=cleandata$Churn), position="dodge") + 
     theme(plot.title = element_text(hjust = 0.5))
print(Category_PhoneService + labs(y="Proportion", x="Phone Service", title="Barplot of Phone Service vs. Churn") +
           scale_fill_discrete(name="Churn"))
```
EDA Barplot of MultipleLines vs. Churn
```{r}
subDF6 <- data.frame(cleandata$MultipleLines, cleandata$Churn)
Category_MultipleLines <- ggplot(subDF6, aes(cleandata$MultipleLines, (..count..)/sum(..count..))) + 
     geom_bar(aes(fill=cleandata$Churn), position="dodge") + 
     theme(plot.title = element_text(hjust = 0.5))
print(Category_MultipleLines + labs(y="Proportion", x="Multiple Lines", title="Barplot of Multiple Lines vs. Churn") +
           scale_fill_discrete(name="Churn"))
```
EDA Barplot of InternetService vs. Churn
```{r}
subDF7 <- data.frame(cleandata$InternetService, cleandata$Churn)
Category_InternetService <- ggplot(subDF7, aes(cleandata$InternetService, (..count..)/sum(..count..))) + 
     geom_bar(aes(fill=cleandata$Churn), position="dodge") + 
     theme(plot.title = element_text(hjust = 0.5))
print(Category_InternetService + labs(y="Proportion", x="Internet Service", title="Barplot of Internet Service vs. Churn") +
           scale_fill_discrete(name="Churn"))
```
EDA Barplot of OnlineSecurity vs. Churn
```{r}
subDF8 <- data.frame(cleandata$OnlineSecurity, cleandata$Churn)
Category_OnlineSecurity <- ggplot(subDF8, aes(cleandata$OnlineSecurity, (..count..)/sum(..count..))) + 
     geom_bar(aes(fill=cleandata$Churn), position="dodge") + 
     theme(plot.title = element_text(hjust = 0.5))
print(Category_OnlineSecurity + labs(y="Proportion", x="Online Security", title="Barplot of Online Security vs. Churn") +
           scale_fill_discrete(name="Churn"))
```
EDA Barplot of OnlineBackup vs. Churn
```{r}
subDF9 <- data.frame(cleandata$OnlineBackup, cleandata$Churn)
Category_OnlineBackup <- ggplot(subDF9, aes(cleandata$OnlineBackup, (..count..)/sum(..count..))) + 
     geom_bar(aes(fill=cleandata$Churn), position="dodge") + 
     theme(plot.title = element_text(hjust = 0.5))
print(Category_OnlineBackup + labs(y="Proportion", x="Online Backup", title="Barplot of Online Backup vs. Churn") +
           scale_fill_discrete(name="Churn"))
```
EDA Barplot of DeviceProtection vs. Churn
```{r}
subDF10 <- data.frame(cleandata$DeviceProtection, cleandata$Churn)
Category_DeviceProtection <- ggplot(subDF10, aes(cleandata$DeviceProtection, (..count..)/sum(..count..))) + 
     geom_bar(aes(fill=cleandata$Churn), position="dodge") + 
     theme(plot.title = element_text(hjust = 0.5))
print(Category_DeviceProtection + labs(y="Proportion", x="Device Protection", title="Barplot of Device Protection vs. Churn") +
           scale_fill_discrete(name="Churn"))
```
EDA Barplot of TechSupport vs. Churn
```{r}
subDF11 <- data.frame(cleandata$TechSupport, cleandata$Churn)
Category_TechSupport <- ggplot(subDF11, aes(cleandata$TechSupport, (..count..)/sum(..count..))) + 
     geom_bar(aes(fill=cleandata$Churn), position="dodge") + 
     theme(plot.title = element_text(hjust = 0.5))
print(Category_TechSupport + labs(y="Proportion", x="Tech Support", title="Barplot of Tech Support vs. Churn") +
           scale_fill_discrete(name="Churn"))
```
EDA Barplot of StreamingTV vs. Churn
```{r}
subDF12 <- data.frame(cleandata$StreamingTV, cleandata$Churn)
Category_StreamingTV <- ggplot(subDF12, aes(cleandata$StreamingTV, (..count..)/sum(..count..))) + 
     geom_bar(aes(fill=cleandata$Churn), position="dodge") + 
     theme(plot.title = element_text(hjust = 0.5))
print(Category_StreamingTV + labs(y="Proportion", x="Streaming TV", title="Barplot of Streaming TV vs. Churn") +
           scale_fill_discrete(name="Churn"))
```
EDA Barplot of StreamingMovies vs. Churn
```{r}
subDF13 <- data.frame(cleandata$StreamingMovies, cleandata$Churn)
Category_StreamingMovies <- ggplot(subDF13, aes(cleandata$StreamingMovies, (..count..)/sum(..count..))) + 
     geom_bar(aes(fill=cleandata$Churn), position="dodge") + 
     theme(plot.title = element_text(hjust = 0.5))
print(Category_StreamingMovies + labs(y="Proportion", x="Streaming Movies", title="Barplot of Streaming Movies vs. Churn") +
           scale_fill_discrete(name="Churn"))
```
EDA Barplot of Contract vs. Churn
```{r}
subDF14 <- data.frame(cleandata$Contract, cleandata$Churn)
Category_Contract <- ggplot(subDF14, aes(cleandata$Contract, (..count..)/sum(..count..))) + 
     geom_bar(aes(fill=cleandata$Churn), position="dodge") + 
     theme(plot.title = element_text(hjust = 0.5))
print(Category_Contract + labs(y="Proportion", x="Contract", title="Barplot of Contract vs. Churn") +
           scale_fill_discrete(name="Churn"))
```
EDA Barplot of PaperlessBilling vs. Churn
```{r}
subDF15 <- data.frame(cleandata$PaperlessBilling, cleandata$Churn)
Category_PaperlessBilling <- ggplot(subDF15, aes(cleandata$PaperlessBilling, (..count..)/sum(..count..))) + 
     geom_bar(aes(fill=cleandata$Churn), position="dodge") + 
     theme(plot.title = element_text(hjust = 0.5))
print(Category_PaperlessBilling + labs(y="Proportion", x="Paperless Billing", title="Barplot of Paperless Billing vs. Churn") +
           scale_fill_discrete(name="Churn"))
```
EDA Barplot of PaymentMethod vs. Churn
```{r}
subDF16 <- data.frame(cleandata$PaymentMethod, cleandata$Churn)
Category_PaymentMethod <- ggplot(subDF16, aes(cleandata$PaymentMethod, (..count..)/sum(..count..))) + 
     geom_bar(aes(fill=cleandata$Churn), position="dodge") + 
     theme(plot.title = element_text(hjust = 0.5))
print(Category_PaymentMethod + labs(y="Proportion", x="Payment Method", title="Barplot of Payment Method vs. Churn") +
           scale_fill_discrete(name="Churn"))
```
EDA Boxplot of Monthly Charges vs. Churn
```{r}
means <- aggregate(MonthlyCharges ~ Churn, cleandata, mean)
Category_Monthly <- ggplot(data=cleandata, aes(x=Churn, y=MonthlyCharges, fill=Churn)) + geom_boxplot() + 
     stat_summary(fun.y=mean, geom="point", shape=15, size=2) +
     geom_text(data=means, aes(label=round(MonthlyCharges,3), y=MonthlyCharges -3.0)) +
     theme(plot.title = element_text(hjust = 0.5))
print(Category_Monthly + labs(y="Monthly Charges ($)", title="Boxplot of Monthly Charges vs. Churn"))
```
EDA Barplot of Customer Tenure (Year)
```{r}
Category_Year <- ggplot(data = cleandata, aes(Year, fill=Year)) + 
     geom_bar(fill="steelblue") +
     geom_text(stat='count', aes(label=..count..), vjust=1.6, color="white", size=3.5) +
     theme_minimal() + 
     theme(plot.title = element_text(hjust = 0.5))
print(Category_Year + labs(x="Tenure (Year)", y="Count", title="Customer Tenure"))
```


***

# 6 Logistic Regression - Preparation 

Make nominal categorical variables to ordinal categorical variables. 
And change some of our categorical variables to binary categorical variables. 
Regard "No Internet Service" as "No", and change the corresponding answers accordingly. 
```{r}
cleandata<-cleandata%>%mutate(gender=ifelse(gender=="Male",1,0))
cleandata<-cleandata%>%mutate(SeniorCitizen=ifelse(SeniorCitizen=="Yes",1,0))
cleandata<-cleandata%>%mutate(Partner=ifelse(Partner=="Yes",1,0))
cleandata<-cleandata%>%mutate(Dependents=ifelse(Dependents=="Yes",1,0))
cleandata<-cleandata%>%mutate(PhoneService=ifelse(PhoneService=="Yes",1,0))
cleandata<-cleandata%>%mutate(MultipleLines=ifelse(MultipleLines=="Yes",1,0))
cleandata<-cleandata%>%mutate(OnlineSecurity=ifelse(OnlineSecurity=="Yes",1,0))
cleandata<-cleandata%>%mutate(OnlineBackup=ifelse(OnlineBackup=="Yes",1,0))
cleandata<-cleandata%>%mutate(DeviceProtection=ifelse(DeviceProtection=="Yes",1,0))
cleandata<-cleandata%>%mutate(TechSupport=ifelse(TechSupport=="Yes",1,0))
cleandata<-cleandata%>%mutate(StreamingTV=ifelse(StreamingTV=="Yes",1,0))
cleandata<-cleandata%>%mutate(StreamingMovies=ifelse(StreamingMovies=="Yes",1,0))
cleandata<-cleandata%>%mutate(PaperlessBilling=ifelse(PaperlessBilling=="Yes",1,0))
cleandata<-cleandata%>%mutate(Churn=ifelse(Churn=="Yes",1,0))
```
Standardize MonthlyCharges variable
```{r}
cleandata$MonthlyCharges <- scale(cleandata$MonthlyCharges)
```
Creating a Baseline model
```{r}
table(cleandata$Churn)/nrow(cleandata)  # Churn rate: 73.42%
```
Split the dataset into training and test sets
```{r warning=FALSE}
set.seed(12345)
sample_size <- floor(0.75 * nrow(cleandata))
sampling_data <- sample(seq_len(nrow(cleandata)), size = sample_size)
train_data <- cleandata[sampling_data, ]
test_data <- cleandata[-sampling_data, ]
```

***

# 7 Logistic Regression

Fit the logistic regression
```{r}
log_reg1 <- glm(Churn ~ ., data=train_data, family=binomial(link="logit"))
summary(log_reg1)
```
Check the deviance
```{r}
anova(object=log_reg1, test="Chisq")
```
Evaluate the logistic regression model
```{r}
predict_train <- predict(log_reg1, newdata=test_data, type="response")
predict_train <- ifelse(predict_train > 0.5, 1, 0)
predict_error <- mean(predict_train != test_data$Churn)
model1 <- 1 - predict_error
print(model1)
```
Variable selection
```{r}
step <- stepAIC(log_reg1, trace=FALSE)
step$anova
```
Summary of our second logistic regression model
```{r}
summary(step)
```
Evaluate the model
```{r}
predict_train2 <- predict(step, newdata=test_data, type="response")
predict_train2 <- ifelse(predict_train2 > 0.5, 1, 0)
predict_error2 <- mean(predict_train2 != test_data$Churn)
model2 <- 1 - predict_error2
print(model2)
```
Confusion matrix
```{r}
table(ActualResult = test_data$Churn, Prediction = predict_train > 0.5)
table(ActualResult = test_data$Churn, Prediction = predict_train2 > 0.5)
```

***

# 8 Conclusion

To sum up, if we use a cutoff point of 0.5 and just guess without any statistical models, 
we will get roughly 73% right. However, we will get approximately 80% right if we use our second 
logistic regression model. And this increased accuracy will help the Telco company predict whether 
a customer will churn or not more accurately, and this analysis will be helpful for them when they 
develop their business strategy to retain their customers. 
